---
title: "TS2_HW1"
author: "Jiwon Lee"
date: "2022-10-10"
output: html_document
---

# Time Series II Homework

```{r}
#download libraries

library(tseries)
library(forecast)
library(haven)
library(fma)
library(expsmooth)
library(lmtest)
library(zoo)
library(seasonal)
library(ggplot2)
library(seasonalview)
library(aTSA)
library(imputeTS)
install.packages("prophet")
library(prophet)
```


## Import Data

```{r}
# original data
hrl_load_metered <- read.csv("~/code/R/Time Series II/TS2_HW1_data/hrl_load_metered.csv")

# dropping unnecessary variables
data <- hrl_load_metered[,c(1,6)]

###### data is hourly from 8/1/19 0:00 to 9/22/22 23:00

# loading test data

test1 <- read.csv("~/code/R/Time Series II/TS2_HW1_data/hrl_load_metered - test1.csv")

val <- test1 ###### test1 is from 9/23/22 to 9/29/22 --> setting this to validation (1 week)

test2 <- read.csv("~/code/R/Time Series II/TS2_HW1_data/hrl_load_metered - test2.csv")

###### test2 is from 9/30/22 to 10/6/22 (1 week)
```

## Creating Time Series Object

```{r}
# create time series object
energy <- ts(data[,2], start = 2019, frequency = 8766) # frequency = 24 hours * 365.25 days in a year

# autoplot
autoplot(energy) +
  ggtitle("Energy Usage") +
  xlab("Time") +
  ylab("Energy")


####### Questions
# 1) how to change the time on the x label to display correctly? 
# 2) how to account for two 0's (obs 14187, 5283)

####### Observations
#series has an annual seasonality - big increase in the summer, slight increase in the winter
#no trend observed

```

## Decomposition

```{r}
#STL decomposition

decomp_stl <- stl(energy, s.window = 7)
plot(decomp_stl)
autoplot(decomp_stl)

####### Comments
#seasonal pattern accounts for most of the variability. Slight upward trend but considering the length of the bar, it does not account for much variability in the data.
```

## Model Building

### Holt Winters 
```{r}
HW <- hw(energy, seasonal = "additive", h = 168)

####### running into error : frequency too high

```

### Unit Root Test for Seasonal Difference

```{r}
# seasonal unit-root testing
energy %>% nsdiffs()

###### 0 --> deterministic solutions
```

### Deterministic Solutions

#### Fourier Transforms

```{r}
plots <- list()
for (i in seq(6)) {
  fit <- auto.arima(energy, xreg = fourier(energy, K = i),
                    seasonal = FALSE, lambda = NULL)
  plots [[i]] <- autoplot(forecast::forecast(fit,
                                             xreg = fourier(energy, K=i, h=168))) + 
    xlab(paste("K=",i, "     BIC=", round(fit[["bic"]],2))) +
    ylab("") + ylim(400, 2500)
}

gridExtra::grid.arrange(
  plots[[1]],plots[[2]],plots[[3]],plots[[4]],plots[[5]],plots[[6]],nrow = 3
)
```


#### Predictor Variables for Seasonality

This approach might be best for our data since the yearly seasonality revolves around the seasons. 
We can have one variable for summer and one for winter.
















