---
title: "TS2_HW1"
author: "Jiwon Lee"
date: "2022-10-10"
output: html_document
---

# Time Series II Homework

```{r}
#download libraries

library(tseries)
library(forecast)
library(haven)
library(fma)
library(expsmooth)
library(lmtest)
library(zoo)
library(seasonal)
library(ggplot2)
library(seasonalview)
library(aTSA)
library(imputeTS)
#install.packages("prophet")
library(prophet)
```


## Import Data

```{r}
# original data
hrl_load_metered <- read.csv("~/Downloads/Homework1_TS2/hrl_load_metered.csv")

# dropping unnecessary variables
data <- hrl_load_metered[,c(1,6)]

for (i in 1:nrow(data)) {
  if (data$mw[i] == 0) {
    data$mw[i] = (data$mw[i + 1] + data$mw[i - 1]) / 2
  }
}

ru_data <- data.frame()
for (i in 1:(nrow(data)/6)) {
  ru_mw <- (1/6) * (data$mw[6 * i - 5] + data$mw[6 * i - 4] + data$mw[6 * i - 3] + data$mw[6 * i - 2] + data$mw[6 * i - 1] + data$mw[6 * i])
  ru_data = rbind(ru_data, c(data$datetime_beginning_ept[4 * i - 3], ru_mw))
}
colnames(ru_data) <- c("Period", "mw")

###### data is hourly from 8/1/19 0:00 to 9/22/22 23:00

# loading test data

test1 <- read.csv("~/Downloads/hrl_load_metered - test1.csv")

val <- test1 ###### test1 is from 9/23/22 to 9/29/22 --> setting this to validation (1 week)

test2 <- read.csv("~/Downloads/hrl_load_metered - test2.csv")

###### test2 is from 9/30/22 to 10/6/22 (1 week)
```

## Creating Time Series Object

```{r}
# create time series object
energy <- ts(data[,2], start = 2019 + 212/365, frequency = 8766) # frequency = 24 hours * 365.25 days in a year
energy_ru <- ts(ru_data[,2], start = 2019 + 212/365, frequency = 1532)
energy_model <- ts(data[,2], start = 2019 + 212/365, frequency = 24)
# autoplot
energy_plot <- autoplot(energy) +
  ggtitle("Hourly Energy Usage from 8/1/19 to 9/22/22") +
  xlab("Time") +
  ylab("Energy (megawatts)")

tseq <- seq.POSIXt(from = as.POSIXct("2019-08-01 00:00:00"), length.out = nrow(data), by = "hours")
# date_vec <- lubridate::seq(from = as.POSIXct("08-01-2019 00:00:00", format = "%m-%d-%Y %h:%m:%s"), to = as.POSIXct("09-22-2022 23:00:00", format = "%m-%d-%y %h:%m:%s"), by = "4 hr")

energy_stl <- stl(energy, s.window = 24)
trend <- ggplot(data = data, aes(x = tseq, y = energy_stl$time.series[,"trend"])) + 
  geom_line() + labs(x = "Time", y = "Energy (megawatts)", title = "Trend of Hourly Energy Usage") + 
  theme(plot.title = element_text(hjust = 0.5))
trend

season <- ggplot(data = data, aes(x = tseq, y = energy_stl$time.series[,"seasonal"])) + 
  geom_line() + labs(x = "Time", y = "Energy (megawatts)", title = "Seasonality of Hourly Energy Usage") + 
  theme(plot.title = element_text(hjust = 0.5))
season

error <- ggplot(data = data, aes(x = tseq, y = energy_stl$time.series[,"remainder"])) + 
  geom_line() + labs(x = "Time", y = "Energy (megawatts)", title = "Error of Hourly Energy Usage") + 
  theme(plot.title = element_text(hjust = 0.5))
error

grid.arrange(energy_plot, grid.arrange(trend, season, error, ncol = 1), ncol = 2)

autoplot(energy_ru) +
  ggtitle("Energy Usage") +
  xlab("Time") +
  ylab("Energy")

####### Questions
# 1) how to change the time on the x label to display correctly? 
# 2) how to account for two 0's (obs 14187, 5283)

####### Observations
#series has an annual seasonality - big increase in the summer, slight increase in the winter
#no trend observed

```

## Decomposition

```{r}
#STL decomposition

# decomp_stl <- stl(energy, s.window = 7)
# plot(decomp_stl)
# autoplot(decomp_stl)

####### Comments
#seasonal pattern accounts for most of the variability. Slight upward trend but considering the length of the bar, it does not account for much variability in the data.
```

## Model Building

### Holt Winters 
```{r}
HW_add <- hw(energy_model, seasonal = "additive", h = 168)
HW_mult <- hw(energy_model, seasonal = "multiplicative", h = 168)

#whole dataset
tseq_with_forecast <- c(tseq, seq.POSIXt(from = as.POSIXct("2022-09-23 00:00:00", tz = "EST"), length.out = 168, by = "hours"))
#only 2022
begin_2022 <- match("1/1/22 0:00", data$datetime_beginning_ept)
tseq_2022 <- seq.POSIXt(from = as.POSIXct("2022-01-01 00:00:00", tz = "EST"), length.out = nrow(data) - begin_2022 + 170, by = "hours")
tseq_2022 <- tseq_2022[c(seq(1, match(as.POSIXct("2022-03-13 01:00:00", tz = "EST"), tseq_2022)), seq(match(as.POSIXct("2022-03-13 03:00:00", tz = "EST"), tseq_2022), length(tseq_2022)))]
#from 8/1/22, same order for each code block
begin_aug_2022 <- match("8/1/22 0:00", data$datetime_beginning_ept)
tseq_2022_aug <- seq.POSIXt(from = as.POSIXct("2022-08-01 00:00:00", tz = "EST"), length.out = nrow(data) - begin_aug_2022 + 169, by = "hours")

#add forecasted values for add
whole_mw_with_forecast_add <- c(data$mw, HW_add$mean)
mw_with_forecast_add <- c(data$mw[begin_2022:nrow(data)], HW_add$mean)
aug_mw_with_forecast_add <- c(data$mw[begin_aug_2022:nrow(data)], HW_add$mean)

#add forecasted values for mult
whole_mw_with_forecast_mult <- c(data$mw, HW_mult$mean)
mw_with_forecast_mult <- c(data$mw[begin_2022:nrow(data)], HW_mult$mean)
aug_mw_with_forecast_mult <- c(data$mw[begin_aug_2022:nrow(data)], HW_mult$mean)

#create dummy df for ggplot
whole_dummy_df <- data.frame(cbind(seq(1:length(tseq_with_forecast)), c(rep(FALSE, (length(tseq))), rep(TRUE, 168))))
dummy_df <- data.frame(cbind(seq(1:length(tseq_2022)), c(rep(FALSE, (length(tseq_2022) - 168)), rep(TRUE, 168))))
dummy_df_aug <- data.frame(cbind(seq(1:length(tseq_2022_aug)), c(rep(FALSE, (length(tseq_2022_aug) - 168)), rep(TRUE, 168))))

#add names
colnames(whole_dummy_df) <- c("index", "forecast")
colnames(dummy_df) <- c("index", "forecast")
colnames(dummy_df_aug) <- c("index", "forecast")

#time plots with forecast colored: add
ggplot(data = whole_dummy_df, aes(x = tseq_with_forecast, y = whole_mw_with_forecast_add, color = factor(forecast), group = 1)) + scale_x_datetime(date_labels = "%Y") + geom_line() + scale_color_manual(values = c("#648fff", "#ffb000"), labels = c("Original", "Forecast")) + labs(x = "Time", y = "Energy (megawatts)", title = "Additive HW Forecast for 9/23/22-9/29/22", color = "Data")
ggplot(data = dummy_df, aes(x = tseq_2022, y = mw_with_forecast_add, color = factor(forecast), group = 1)) + scale_x_datetime(date_labels = "%m/%Y") + geom_line() + scale_color_manual(values = c("#648fff", "#ffb000"), labels = c("Original", "Forecast")) + labs(x = "Time", y = "Energy (megawatts)", title = "Additive HW Forecast for 9/23/22-9/29/22", color = "Data")
ggplot(data = dummy_df_aug, aes(x = tseq_2022_aug, y = aug_mw_with_forecast_add, color = factor(forecast), group = 1)) + scale_x_datetime(date_labels = "%m/%d/%Y") + geom_line() + scale_color_manual(values = c("#648fff", "#ffb000"), labels = c("Original", "Forecast")) + labs(x = "Time", y = "Energy (megawatts)", title = "Additive HW Forecast for 9/23/22-9/29/22", color = "Data")

#time plots with forecast colored: mult
ggplot(data = whole_dummy_df, aes(x = tseq_with_forecast, y = whole_mw_with_forecast_mult, color = factor(forecast), group = 1)) + scale_x_datetime(date_labels = "%Y") + geom_line() + scale_color_manual(values = c("#648fff", "#ffb000"), labels = c("Original", "Forecast")) + labs(x = "Time", y = "Energy (megawatts)", title = "Multiplicative HW Forecast for 9/23/22-9/29/22", color = "Data")
ggplot(data = dummy_df, aes(x = tseq_2022, y = mw_with_forecast_mult, color = factor(forecast), group = 1)) + scale_x_datetime(date_labels = "%m/%Y") + geom_line() + scale_color_manual(values = c("#648fff", "#ffb000"), labels = c("Original", "Forecast")) + labs(x = "Time", y = "Energy (megawatts)", title = "Multiplicative HW Forecast for 9/23/22-9/29/22", color = "Data")
ggplot(data = dummy_df_aug, aes(x = tseq_2022_aug, y = aug_mw_with_forecast_mult, color = factor(forecast), group = 1)) + scale_x_datetime(date_labels = "%m/%d/%Y") + geom_line() + scale_color_manual(values = c("#648fff", "#ffb000"), labels = c("Original", "Forecast")) + labs(x = "Time", y = "Energy (megawatts)", title = "Multiplicative HW Forecast for 9/23/22-9/29/22", color = "Data")

add_sum = 0
mult_sum = 0
for (i in 1:nrow(val)) {
  add_sum <- add_sum + (abs(val$mw[i] - HW_add$mean[i]) / val$mw[i])
  mult_sum <- mult_sum + (abs(val$mw[i] - HW_mult$mean[i]) / val$mw[i])
}
add_mape = add_sum / nrow(val)
mult_mape = mult_sum / nrow(val)

#summary(HW)
####### running into error : frequency too high

```

### Unit Root Test for Seasonal Difference

```{r}
# seasonal unit-root testing
energy_model %>% nsdiffs()
energy_model %>% diff(lag = 24) %>% ggtsdisplay()

###### 0 --> deterministic solutions
```

### Stochastic Solutions

#### Seasonal ARIMA

```{r}

first_model = 
second_model = Arima(energy_model, order = c(2, 0, 1), seasonal = c(3, 1, 0))
second_model %>% residuals() %>% ggtsdisplay() + labs(title = "Second Model: ARIMA(2,0,1)(3,1,0)[24]")
auto_model = auto.arima(energy_model, seasonal = TRUE)
auto_model %>% residuals() %>% ggtsdisplay() + labs(title = "Auto Model: ARIMA(5,0,0)(2,1,0)[24]")

```


#### Predictor Variables for Seasonality

This approach might be best for our data since the yearly seasonality revolves around the seasons. 
We can have one variable for summer and one for winter.
















